05_DESeq_manipulation

### 2019-05-19 analysis/05_DESeq2_manipulate.Rmd
now with many loops and functions

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)
library(edgeR)

@@

## Create variance stabilized data

```{r vsd}

mydds <- list(ddsmale_hypothalamus) # for testing only
#mydds <- list(ddsfemale_hypothalamus, ddsfemale_pituitary, ddsfemale_gondad,
#              ddsmale_hypothalamus, ddsmale_pituitary, ddsmale_gondad)

  vdsname <- paste("vsd", eachgroup, sep = "")
  assign(paste("vsd", eachgroup, sep = ""), vsd)
  return(vdsname)
}

@@

```{r totalDEGs}

@@

  head(totalDEGS)

  # widen data to create table of degs
  rownames(totalDEGS) <- NULL #remove row names
  totalDEGS_wide <- spread(totalDEGS, V2, V3)
  print(totalDEGS_wide)

  totalDEGS$V1 <- factor(totalDEGS$V1, levels = 
                              c("m.inc.d3",  "m.inc.d8",
                                "m.inc.d9", "m.inc.d17",
                                "prolong", "extend", "m.n2"))
  totalDEGS$V2 <- factor(totalDEGS$V2, levels = 
                              c("m.inc.d3",  "m.inc.d8",
                                "m.inc.d9", "m.inc.d17",
                                "prolong", "extend", "m.n2"))

  allcontrasts <- totalDEGS %>%
    ggplot( aes(V1, V2)) +
      geom_tile(aes(fill = V3)) +
      scale_fill_viridis(na.value="#440154", 
                      limits = c(0, 3000),
                      breaks = c(0, 1000, 2000, 3000)) + 
      xlab(" ") + ylab("Timepoint") +
      labs(fill = "# of DEGs",
          subtitle = "eachgroup")
  plot(allcontrasts)
}
```

@@

plot_grid(pca1, pca2)

@@

@@

# add pvalues

for (i in a){
  for (j in b){
    if (i != j) {
      k <- paste(i,j, sep = " vs ")
      print(k)
      results <- returnpadj(i,j)
      DEGs <- cbind(DEGs,results)
    }
  }
}

@@

pheatmap(sigDEGs, kmeans_k = 5,
         show_rownames = F, show_colnames = F,
         color = viridis(30),
         breaks=myBreaks,
         annotation_col=anndf,
         main = eachgroup)

## candidate genes
names(geneinfo)
names(geneinfo)[4] <- "rownames"
DEGs$rownames <- row.names(DEGs)

@@

p1 <- candidatecounts %>%
  filter(gene %in% c( "AR", "CYP19A1", "ESR1", "ESR2", "FSHR")) %>%
  ggplot(aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot() +
  facet_wrap(~gene, scales = "free", nrow = 1) +
  theme(axis.text.x = element_blank(),
        legend.position = "bottom") +
  labs(subtitle = eachgroup)
p1

print(p1)

cat("\n")

p2 <- candidatecounts %>%
  filter(gene %in% c("GHRL", "GAL", "NPVF", "GNRH1", "LHCGR")) %>%
  ggplot(aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot() +
  facet_wrap(~gene, scales = "free", nrow = 1) +
  theme(axis.text.x = element_blank(),
        legend.position = "bottom") +
  labs(subtitle = eachgroup)
p2

print(p2)

cat("\n")

p3 <- candidatecounts %>%
  filter(gene %in% c("PGR", "PRL", "PRLR", "VIP", "VIPR1")) %>%
  ggplot(aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot() +
  facet_wrap(~gene, scales = "free", nrow = 1) +
  theme(axis.text.x = element_blank(),
        legend.position = "bottom") +
  labs(subtitle = eachgroup)
p3

print(p3)

cat("\n")

cat("Done with eachgroup")

cat("\n")

>>>>

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)

@@

@@

```{r totalDEGs}
#mydds <- list(ddsmale_hypothalamus) # for testing only
mydds <- list(ddsfemale_hypothalamus, ddsfemale_pituitary, ddsfemale_gondad,
              ddsmale_hypothalamus, ddsmale_pituitary, ddsmale_gondad)

@@

  totalDEGSname <- paste("totalDEGS", eachgroup, sep = "")
  assign(paste("totalDEGS", eachgroup, sep = ""), totalDEGS)
  return(totalDEGSname)
}

a <- plottotalDEGs(totalDEGSfemale_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(totalDEGSfefemale_pituitary, "female pituitary")
c <- plottotalDEGs(totalDEGSmale_gonad, "female gonad")
d <- plottotalDEGs(totalDEGSmale_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(totalDEGSmale_pituitary, "male pituitary")
f <- plottotalDEGs(totalDEGSmale_gonad, "male gonad")

plot_grid(a,b,c,d,e,f, nrow = 2)
```

## Calculate and plot principal components

```{r pca}

@@

  mypca <- plot_grid(pca1, pca2)
  plot(mypca)

@@

```{r pheatmap}

@@

  a <- levels(colData$treatment)
  b <- levels(colData$treatment)

  for (i in a){
    for (j in b){
      if (i != j) {
        print(i)
        results <- returnpadj(i,j)
        DEGs <- cbind(DEGs,results)
      }
    }
     b <- b[-1]  # drop 1st element of second string to not recalculate DEGs
  }

@@

  pheatmap(sigDEGs, kmeans_k = 5,
         show_rownames = F, show_colnames = F,
         color = viridis(30),
         breaks=myBreaks,
         annotation_col=anndf,
         main = eachgroup)
}

```{r candidates}
for (eachdds in mydds){

names(geneinfo)
names(geneinfo)[4] <- "rownames"
DEGs$rownames <- row.names(DEGs)

@@

p1 <- candidatecounts %>%
  filter(gene %in% c( "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
                      "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
                      "PGR", "PRL", "PRLR", "VIP", "VIPR1")) %>%
  ggplot(aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot() +
  facet_wrap(~gene, scales = "free") +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_blank(),
        legend.position = "bottom") +
  labs(subtitle = eachgroup, x = NULL) +
  guides(fill= guide_legend(nrow=1))
p1
